---
title: "Multiple Regression Notes"
author: "D. Spence"
date: "November 3, 2016"
output: html_document
---

```{r Preliminaries, results="hide", message=FALSE, warning=FALSE}
library("ggplot2")
data("mtcars")
data("InsectSprays")
```

## Categorical Predictors

### A Simple Categorical Variable
```{r SimpleCategorical}
amcode <- function(x) {if (x) '1-Manual' else '2-Automatic'}
mtcars$trans <- sapply(mtcars$am,amcode)

boxplot(mpg ~ trans, data = mtcars, xlab="Transmission", ylab="MPG")

transModel <- lm(mpg ~ trans, data=mtcars)
summary(transModel)

a <- transModel$coefficients[1]
b <- transModel$coefficients[2]
```

$\widehat{MPG} = `r round(a,2)` + `r round(b,2)`*trans$  
For manual transmission, trans is 0. This model indicates the mean gas mileage for cars with manual transmission is `r round(a, 1)` MPG. For automatic transmission, trans is 1. This model indicates the mean gas mileage for cars with automatic transmission is `r round(a+b,1)` MPG.

### Categorical Variable with Several Values
```{r SeveralCategories}
boxplot(count ~ spray, data = InsectSprays, xlab="Spray", ylab="Insects Found")
sprayModel <- lm(count ~ spray, data = InsectSprays)

summary(sprayModel)
```

## More Than One Predictor

```{r TwoPredictors}
mpgModel <- lm(mpg ~ wt + trans, data = mtcars)

summary(mpgModel)
```

```{r SeveralPredictors}
mpgModel <- lm(mpg ~ ., data = mtcars)

summary(mpgModel)
```
